---
title: "cum_heat_calculation"
output: html_document
date: "2023-02-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Cumulative Heat Calculation

For each cherry blossom location, we calculated the cumulative heat by adding daily maximum temperature (in celcius) to the date-of-year(doy) of cherry blossom. Considering the difference in local climate, we will need to define start date to add cumulative heat and cumulative heat threshold. 

We used exhaustive approach to estimate start date and threshold by adding daily max temperature with start date ranging from first day of year to the 70_th day of year. To estimate the optimal start date and threshold temperature, we subset the last ten years of temperature data to check accuracy of our estimate. The start date with the lowest absolute value error is selected.

```{r}
cum.heat.standard <- data.frame()
```


#Liestal

## Obtain necessary data

```{r}
weather <- read.csv("../../data/liestal/liestal.csv")
bloom <- read.csv("../../data/liestal/liestaldoy.csv")

cum.heat <- data.frame(bloom$year, bloom$bloom_doy)
colnames(cum.heat) <- c("year", "doy")
```


## Add day of year to weather (for future calculation)

```{r}
current.year <- 0
count <- 1
weather.doy <- c()
for(i in 1:nrow(weather)) {
  this.year <- weather[i, 1] %/% 10000
  if (this.year != current.year) {  # if year changed
    current.year <- this.year
    count <- 1
  } else {
    count <- count + 1
  }
  weather.doy <- append(weather.doy, count)
}

weather <- cbind(weather, weather.doy)
```

## conver max temp from Fahrenheit to celcuis

```{r}
max.celcuis <- (weather$Max.Temp - 32) * (5/9)
weather <- cbind(weather, max.celcuis)
```


## List of potential start dates for cumulative heat calculation

```{r}
start.doy <- c(1:70)
```


## Create data frame containing calculation of cumulative heat from all potential start dates

```{r}
cumulative.heat <- c()
dynamic.doy <- data.frame(bloom$year)

for (startDate in start.doy) {
  for (year in cum.heat$year) {
    if (weather$Date[1] <= year* 10000 + 101) {
      cumHeat <- 0
      daily.temp <- weather[which(weather$Date %/% 10000 == year), c("max.celcuis")]
      for(i in startDate:bloom[which(bloom$year == year), c("bloom_doy")]) {
        cumHeat = cumHeat + daily.temp[i]
      }
      cumulative.heat <- append(cumulative.heat, cumHeat)
    } else {
      cumulative.heat <- append(cumulative.heat, NA)
    }
  }
  dynamic.doy <- cbind(dynamic.doy, cumulative.heat)
  cumulative.heat <- c()
}
```

## name columns and remove years without temperature record

```{r}
colnames(dynamic.doy) <- c("year", start.doy)

dynamic.doy <- na.omit(dynamic.doy)
```


## separate a testing set

```{r}
train <- dynamic.doy[c(1:(nrow(dynamic.doy)-10)), ]
test <- tail(dynamic.doy, n = 10)
```

## Obtain mean cumulative heat for each start date in dynamic.doy

```{r}
mean.heat <- c()
for(col in 2:ncol(train)) {
  mean.heat <- append(mean.heat, mean(train[, col]))
}
```

```{r}
yearly.prediction <- data.frame()

for (year in test$year) {
  max.temps <- weather[which(weather$Date %/% 10000 == year) ,]$max.celcuis
  predict.doys <- c()
  for (startDoy in start.doy) {
    target.heat <- mean.heat[startDoy]
    cumHeat <- 0
    predicted.doy <- startDoy
    while (cumHeat < target.heat) {
      predicted.doy = predicted.doy + 1
      cumHeat = cumHeat + max.temps[predicted.doy]
    }
    predict.doys <- append(predict.doys, predicted.doy)
  }
  yearly.prediction <- rbind(yearly.prediction, predict.doys)
}
yearly.prediction <- cbind(test$year, yearly.prediction)
colnames(yearly.prediction) <- c("year", start.doy)
yearly.prediction
```

## Calculate error between prediction and actual day of blossom

```{r}
yearly.error.abs <- data.frame()
true.doy <- tail(bloom, n = 10)$bloom_doy
for (year in 1:length(test$year)){
  error <- c()
  for (i in start.doy) {
    error <- append(error, abs(true.doy[year] - yearly.prediction[year, i + 1]))
  }
  yearly.error.abs <- rbind(yearly.error.abs, error)
}

yearly.error.abs <- cbind(test$year, yearly.error.abs)
colnames(yearly.error.abs) <- c("year", start.doy)
yearly.error.abs
```

## mean error for each selected start date & optimal start doy with minimal mean error

```{r}
mean.abs.error <- c()
for (col in 2:ncol(yearly.error.abs)) {
  mean.abs.error <- append(mean.abs.error, mean(yearly.error.abs[, col]))
}

mean.abs.error <- data.frame(start.doy, mean.abs.error)
mean.abs.error
best.startdoy <- mean.abs.error[which.min(mean.abs.error$mean.abs.error), ]$start.doy
best.startdoy
avg.threshold <- mean(dynamic.doy[, best.startdoy + 1])

cum.heat.standard <- rbind(cum.heat.standard, c(best.startdoy, avg.threshold))
```


```{r}
optimal.doy <- mean.abs.error[which.min(mean.abs.error$mean.abs.error), ]$start.doy
temp.threshold <- dynamic.doy[, (optimal.doy + 1)]

kyoto.cumulative <- data.frame(dynamic.doy$year, temp.threshold)
colnames(kyoto.cumulative) <- c("year", "cumulative.temp")
write.csv(kyoto.cumulative, "../../data/liestal/liestal_cumTemp.csv")
```


# Washington DC

## Obtain necessary data

```{r}
weather <- read.csv("../../data/washington/washington.csv")
bloom <- read.csv("../../data/washington/washingtondoy.csv")

cum.heat <- data.frame(bloom$year, bloom$bloom_doy)
colnames(cum.heat) <- c("year", "doy")
```


## Add day of year to weather (for future calculation)

```{r}
current.year <- 0
count <- 1
weather.doy <- c()
for(i in 1:nrow(weather)) {
  this.year <- weather[i, 1] %/% 10000
  if (this.year != current.year) {  # if year changed
    current.year <- this.year
    count <- 1
  } else {
    count <- count + 1
  }
  weather.doy <- append(weather.doy, count)
}

weather <- cbind(weather, weather.doy)
```

## conver max temp from Fahrenheit to celcuis

```{r}
max.celcuis <- (weather$Max.Temp - 32) * (5/9)
weather <- cbind(weather, max.celcuis)
```


## List of potential start dates for cumulative heat calculation

```{r}
start.doy <- c(1:70)
```


## Create data frame containing calculation of cumulative heat from all potential start dates

```{r}
cumulative.heat <- c()
dynamic.doy <- data.frame(bloom$year)

for (startDate in start.doy) {
  for (year in cum.heat$year) {
    if (weather$Date[1] <= year* 10000 + 101) {
      cumHeat <- 0
      daily.temp <- weather[which(weather$Date %/% 10000 == year), c("max.celcuis")]
      for(i in startDate:bloom[which(bloom$year == year), c("bloom_doy")]) {
        cumHeat = cumHeat + daily.temp[i]
      }
      cumulative.heat <- append(cumulative.heat, cumHeat)
    } else {
      cumulative.heat <- append(cumulative.heat, NA)
    }
  }
  dynamic.doy <- cbind(dynamic.doy, cumulative.heat)
  cumulative.heat <- c()
}
```

## name columns and remove years without temperature record

```{r}
colnames(dynamic.doy) <- c("year", start.doy)

dynamic.doy <- na.omit(dynamic.doy)
```


## separate a testing set

```{r}
train <- dynamic.doy[c(1:(nrow(dynamic.doy)-10)), ]
test <- tail(dynamic.doy, n = 10)
```

## Obtain mean cumulative heat for each start date in dynamic.doy

```{r}
mean.heat <- c()
for(col in 2:ncol(train)) {
  mean.heat <- append(mean.heat, mean(train[, col]))
}
```

```{r}
yearly.prediction <- data.frame()

for (year in test$year) {
  max.temps <- weather[which(weather$Date %/% 10000 == year) ,]$max.celcuis
  predict.doys <- c()
  for (startDoy in start.doy) {
    target.heat <- mean.heat[startDoy]
    cumHeat <- 0
    predicted.doy <- startDoy
    while (cumHeat < target.heat) {
      predicted.doy = predicted.doy + 1
      cumHeat = cumHeat + max.temps[predicted.doy]
    }
    predict.doys <- append(predict.doys, predicted.doy)
  }
  yearly.prediction <- rbind(yearly.prediction, predict.doys)
}
yearly.prediction <- cbind(test$year, yearly.prediction)
colnames(yearly.prediction) <- c("year", start.doy)
yearly.prediction
```

## Calculate error between prediction and actual day of blossom

```{r}
yearly.error.abs <- data.frame()
true.doy <- tail(bloom, n = 10)$bloom_doy
for (year in 1:length(test$year)){
  error <- c()
  for (i in start.doy) {
    error <- append(error, abs(true.doy[year] - yearly.prediction[year, i + 1]))
  }
  yearly.error.abs <- rbind(yearly.error.abs, error)
}

yearly.error.abs <- cbind(test$year, yearly.error.abs)
colnames(yearly.error.abs) <- c("year", start.doy)
yearly.error.abs
```

## mean error for each selected start date & optimal start doy with minimal mean error

```{r}
mean.abs.error <- c()
for (col in 2:ncol(yearly.error.abs)) {
  mean.abs.error <- append(mean.abs.error, mean(yearly.error.abs[, col]))
}

mean.abs.error <- data.frame(start.doy, mean.abs.error)
mean.abs.error

best.startdoy <- mean.abs.error[which.min(mean.abs.error$mean.abs.error), ]$start.doy
best.startdoy
avg.threshold <- mean(dynamic.doy[, best.startdoy + 1])

cum.heat.standard <- rbind(cum.heat.standard, c(best.startdoy, avg.threshold))
```


```{r}
optimal.doy <- mean.abs.error[which.min(mean.abs.error$mean.abs.error), ]$start.doy
temp.threshold <- dynamic.doy[, (optimal.doy + 1)]

washington.cumulative <- data.frame(dynamic.doy$year, temp.threshold)
colnames(washington.cumulative) <- c("year", "cumulative.temp")
write.csv(washington.cumulative, "../../data/washington/washington_cumTemp.csv")
```


# Kyoto 

## Obtain necessary data

```{r}
weather <- read.csv("../../data/kyoto/kyoto.csv")
bloom <- read.csv("../../data/kyoto/kyotodoy.csv")

cum.heat <- data.frame(bloom$year, bloom$bloom_doy)
colnames(cum.heat) <- c("year", "doy")
```


## Add day of year to weather (for future calculation)

```{r}
current.year <- 0
count <- 1
weather.doy <- c()
for(i in 1:nrow(weather)) {
  this.year <- weather[i, 1] %/% 10000
  if (this.year != current.year) {  # if year changed
    current.year <- this.year
    count <- 1
  } else {
    count <- count + 1
  }
  weather.doy <- append(weather.doy, count)
}

weather <- cbind(weather, weather.doy)
```

## conver max temp from Fahrenheit to celcuis

```{r}
max.celcuis <- (weather$Max.Temp - 32) * (5/9)
weather <- cbind(weather, max.celcuis)
```


## List of potential start dates for cumulative heat calculation

```{r}
start.doy <- c(1:70)
```


## Create data frame containing calculation of cumulative heat from all potential start dates

```{r}
cumulative.heat <- c()
dynamic.doy <- data.frame(bloom$year)

for (startDate in start.doy) {
  for (year in cum.heat$year) {
    if (weather$Date[1] <= year* 10000 + 101) {
      cumHeat <- 0
      daily.temp <- weather[which(weather$Date %/% 10000 == year), c("max.celcuis")]
      for(i in startDate:bloom[which(bloom$year == year), c("bloom_doy")]) {
        cumHeat = cumHeat + daily.temp[i]
      }
      cumulative.heat <- append(cumulative.heat, cumHeat)
    } else {
      cumulative.heat <- append(cumulative.heat, NA)
    }
  }
  dynamic.doy <- cbind(dynamic.doy, cumulative.heat)
  cumulative.heat <- c()
}
```

## name columns and remove years without temperature record

```{r}
colnames(dynamic.doy) <- c("year", start.doy)

dynamic.doy <- na.omit(dynamic.doy)
```


## separate a testing set

```{r}
train <- dynamic.doy[c(1:(nrow(dynamic.doy)-10)), ]
test <- tail(dynamic.doy, n = 10)
```

## Obtain mean cumulative heat for each start date in dynamic.doy

```{r}
mean.heat <- c()
for(col in 2:ncol(train)) {
  mean.heat <- append(mean.heat, mean(train[, col]))
}
```

```{r}
yearly.prediction <- data.frame()

for (year in test$year) {
  max.temps <- weather[which(weather$Date %/% 10000 == year) ,]$max.celcuis
  predict.doys <- c()
  for (startDoy in start.doy) {
    target.heat <- mean.heat[startDoy]
    cumHeat <- 0
    predicted.doy <- startDoy
    while (cumHeat < target.heat) {
      predicted.doy = predicted.doy + 1
      cumHeat = cumHeat + max.temps[predicted.doy]
    }
    predict.doys <- append(predict.doys, predicted.doy)
  }
  yearly.prediction <- rbind(yearly.prediction, predict.doys)
}
yearly.prediction <- cbind(test$year, yearly.prediction)
colnames(yearly.prediction) <- c("year", start.doy)
yearly.prediction
```

## Calculate error between prediction and actual day of blossom

```{r}
yearly.error.abs <- data.frame()
true.doy <- tail(bloom, n = 10)$bloom_doy
for (year in 1:length(test$year)){
  error <- c()
  for (i in start.doy) {
    error <- append(error, abs(true.doy[year] - yearly.prediction[year, i + 1]))
  }
  yearly.error.abs <- rbind(yearly.error.abs, error)
}

yearly.error.abs <- cbind(test$year, yearly.error.abs)
colnames(yearly.error.abs) <- c("year", start.doy)
yearly.error.abs
```

## mean error for each selected start date & optimal start doy with minimal mean error

```{r}
mean.abs.error <- c()
for (col in 2:ncol(yearly.error.abs)) {
  mean.abs.error <- append(mean.abs.error, mean(yearly.error.abs[, col]))
}

mean.abs.error <- data.frame(start.doy, mean.abs.error)
mean.abs.error
best.startdoy <- mean.abs.error[which.min(mean.abs.error$mean.abs.error), ]$start.doy
best.startdoy
avg.threshold <- mean(dynamic.doy[, best.startdoy + 1])

cum.heat.standard <- rbind(cum.heat.standard, c(best.startdoy, avg.threshold))
```


```{r}
optimal.doy <- mean.abs.error[which.min(mean.abs.error$mean.abs.error), ]$start.doy
temp.threshold <- dynamic.doy[, (optimal.doy + 1)]

kyoto.cumulative <- data.frame(dynamic.doy$year, temp.threshold)
colnames(kyoto.cumulative) <- c("year", "cumulative.temp")
write.csv(kyoto.cumulative, "../../data/kyoto/kyoto_cumTemp.csv")
```


# Result

```{r}
cum.heat.standard <- cbind(c("Liestal", "Washington", "Kyoto"), cum.heat.standard)

colnames(cum.heat.standard) <- c("Region", "Start Date", "Threshold_C")

write.csv(cum.heat.standard, "../../data/cumulative_heat_standard.csv")
```

