---
title: "Linear Regression"
output: html_document
date: "2023-02-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Washington

### Linear Regression
```{r}
df1 = read.csv("../../data/washington/washington_clean_combined.csv")
df1 = df1[-2]
set.seed(3)
test = sample(c(1:77), 15, replace = F)
df_test = df1[df1$X %in% test,]
df_train = df1[!(df1$X %in% df_test$X),]
df_test = df_test[-1]
df_train = df_train[-1]
```

```{r}
mod1 = lm(bloom_doy ~ ., data = df_train)
test_predicted = round(predict.lm(mod1, newdata = df_test))
test_real = df_test["bloom_doy"]
residual = abs(test_predicted - test_real)
len = length(residual$bloom_doy)
accNum = length(residual$bloom_doy[residual$bloom_doy <= 3] )
accuracy = accNum / len; accuracy
```

**Linear Regression Accuracy on Testing set: 0.3333333**

### Subset Selection

```{r}
library(caret)
library(leaps)
library(tidyverse)

mod2 = regsubsets(bloom_doy ~ ., data = df_train, nvmax = 5)
summary(mod2)

draw <- function(reg.summary, regfit){
  par(mfrow=c(1,2))
  plot(reg.summary$adjr2,xlab="Number of Variables",ylab="adjusted Rˆ2",type="l") 
  best.mod.ind = which.min(reg.summary$cp) # which size of model gives the best Cp?
  points(best.mod.ind,reg.summary$adjr2[best.mod.ind], col="red",cex=2,pch=20)
  plot(reg.summary$cp,xlab="Number of Variables",ylab="Cp",type="l")
  points(best.mod.ind,reg.summary$cp[best.mod.ind], col="red",cex=2,pch=20)
  plot(reg.summary$bic,xlab="Number of Variables",ylab="BIC",type="l")
  points(best.mod.ind,reg.summary$bic[best.mod.ind], col="red",cex=2,pch=20)
  plot(regfit,scale="bic")
  coef(regfit,best.mod.ind)
}
draw(summary(mod2), mod2)
```

```{r}
predict.regsubsets <- function(object, newdata, id, ...) {
  form <- as.formula(object[["call"]][[2]])
  mat <- model.matrix(form, newdata)
  coefi <- coef(object, id = id)
  xvars <- names(coefi)
  # Matrix
  pred_mat <- mat[, xvars] %*% coefi
  # Vector
  pred <- as.numeric(pred_mat)
  names(pred) <- rownames(mat)
  pred
}

subPred = predict(mod2, newdata = df_test, 3)
resSub = abs(subPred - test_real)
accNum = length(which(resSub <= 3))
accSub = accNum / len; accSub
```
**Subset Selection Accuracy on Testing set: 0.4**

\newpage

# Kyoto

### Linear Regression

```{r}
df2 = read.csv("../../data/kyoto/kyotodoy_yearly.csv")
df2 = df2[-2]
set.seed(3)
test_kyoto = sample(c(1:26), 5, replace = F)
df_test_kyoto = df2[df2$X %in% test_kyoto,]
df_train_kyoto = df2[!(df2$X %in% df_test_kyoto$X),]
df_test_kyoto = df_test_kyoto[-1]
df_train_kyoto = df_train_kyoto[-1]
```

```{r}
mod3 = lm(bloom_doy ~ ., data = df_train_kyoto)
test_predicted_kyoto = round(predict.lm(mod3, newdata = df_test_kyoto))
test_real_kyoto = df_test_kyoto["bloom_doy"]
residual_kyoto = abs(test_predicted_kyoto - test_real_kyoto)
len_kyoto = length(residual$bloom_doy)
accNum_kyoto = length(residual$bloom_doy[residual$bloom_doy <= 3] )
accuracy_kyoto = accNum_kyoto / len_kyoto; accuracy_kyoto
```

**Linear Regression Accuracy on Testing set: 0.3333333**

### Subset Selection
```{r}
mod4 = regsubsets(bloom_doy ~ ., data = df_train_kyoto, nvmax = 6)
summary(mod4)

draw(summary(mod4), mod4)
```

```{r}
subPred_kyoto = predict(mod4, newdata = df_test_kyoto, 3)
resSub_kyoto = abs(subPred_kyoto - test_real_kyoto)
accNum_kyoto = length(which(resSub_kyoto <= 3))
accSub_kyoto = accNum_kyoto / len_kyoto; accSub_kyoto
```

**Subset Selection Accuracy on Testing set: 0.2**

\newpage

# liestal

### Linear Regression

```{r}
df3 = read.csv("../../data/liestal/liestal_yearly.csv")
df3 = df3[-2]
set.seed(3)
test_liestal = sample(c(1:26), 5, replace = F)
df_test_liestal = df3[df3$X %in% test_liestal,]
df_train_liestal = df3[!(df3$X %in% df_test_liestal$X),]
df_test_liestal = df_test_liestal[-1]
df_train_liestal = df_train_liestal[-1]
```

```{r warning=FALSE}
mod5 = lm(bloom_doy ~ ., data = df_train_liestal)
test_predicted_liestal = round(predict.lm(mod5, newdata = df_test_liestal))
test_real_liestal = df_test_liestal["bloom_doy"]
residual_liestal = abs(test_predicted_liestal - test_real_liestal)
len_liestal = length(residual$bloom_doy)
accNum_liestal = length(residual$bloom_doy[residual$bloom_doy <= 3] )
accuracy_liestal = accNum_liestal / len_liestal; accuracy_liestal
```

**Linear Regression Accuracy on Testing set: 0.3333333**

### Subset Selection

```{r}
mod6 = regsubsets(bloom_doy ~ ., data = df_train_liestal, nvmax = 6)
summary(mod6)

draw(summary(mod6), mod6)
```

```{r}
subPred_liestal = predict(mod6, newdata = df_test_liestal, 3)
resSub_liestal = abs(subPred_liestal - test_real_liestal)
accNum_liestal = length(which(resSub_liestal <= 3))
accSub_liestal = accNum_liestal / len_liestal; accSub_liestal
```

**Subset Selection Accuracy on Testing set: 0.1333333**